---
title: "Replication_Mattie_Soghikian.qmd"
format: html
editor: visual
---

# Activity Objectives

Exposure to Bioinformatics work.

## BioInformatics Objectives:

-   Use the arrow (`<-`) operator to store objects in variables.

-   Use R\'s `sample()` function to create a random \"genome\" for the purpose of testing functions.

-   Build functions to process and analyze genetic data.

-   Learn about `for` loops and implement them when necessary.

-   Read in a real bacterial genome from a text file.

-   Analyse the genome using the ideas and tools developed in the first part of the notebook.

## Loading Packages

```{r setup, include = FALSE}
#Load in any packages you need
library(tidyverse)

#Read in any data
```

### Challenge 1

Create a list of nucleotides using `c()`

```{r}

nucleotides <- c("A","C","G","T")
```

### Challenge 2

Generating a random string of 15 nucleotides

```{r}

#Generating a random genome

numNucleotides <- 15

randGenome <- sample(nucleotides, size = numNucleotides, replace = TRUE)

randGenome

#Collapse genome into single string

paste(randGenome, collapse = "")
```

### Challenge 3

Generate random genome 1500 nucleotides long and use `paste()` method to collapse into a single string

```{r}

set.seed(215)

numNucleotides <- 1500

randGenome <- sample(nucleotides, size = numNucleotides, replace = TRUE)

randGenome

#Collapse genome into single string

randGenome <- paste(randGenome, collapse = "")

randGenome
```

### Challenge 4

Set a seed and generate random genome consisting of 100 nucleotides and collapse it.

```{r}
set.seed(215)

numNucleotides <- 100

randGenome <- sample(nucleotides, size = numNucleotides, replace = TRUE)

randGenome

#Collapse genome into single string

randGenome <- paste(randGenome, collapse = "")

randGenome
```

### Challenge 5

Create loop with variable `myProduct` with the value 1. Write your own for loop which takes the integer values from 1-15.

```{r}

myProduct <- 1

for(j in 1:15){
  mySum <- mySum + j
  print(mySum)
}
```

### Challenge 6

Generate a random genome consisting of 10 nucleotides and collapse into a single string rather than a list. Then write a `for()` loop.

```{r}

numNucleotides <- 10

randGenome <- sample(nucleotides, size = numNucleotides, replace = TRUE)

randGenome<- paste(randGenome, collapse = "")

randGenome

for(i in 1:10){
  print(str_sub(randGenome, start=i, end= i))
}
#Flow control - if, else if and else statements
#Generate new random genome

nucleotides <- c("A", "C", "B", "T")
genomeLength <- 10

randGenome <- paste(
  sample(nucleotides, size = genomeLength, replace = TRUE),
  collapse = "")

print(randGenome)

```

### Challenge 7

Adapt the loop so it counts the number of occurences of Adenine in randGenome.

```{r}

randGenome

adenine_count <- 0

#nchar = number of characters

for(i in 1:nchar(randGenome)){
  if(str_sub(randGenome, start = i, end = i) == "A"){
    adenine_count <- adenine_count + 1
  }
}

```
